Technical Report

### 2018-03-21 fully-abstract-comp/technical-report.tex
Change old target macro to be a macro for constructs common between the two languages (that's how it's used) and add a real target macro.
Define initial execution configurations.

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \ints \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}} \mid \\ 
      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
\end{array}
\]
$i$ defines the set $\Instr$

@@

The source machine is also a capability machine with memory capabilities and sealed capabilities. Unlike the target machine, the source machine has a built in stack along with special stack and return tokens used in place of the actual capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    o & \in & \nats \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\offpc,\offsigma}{r}{r}
  \end{array}
\]
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{\offpc,\offsigma}{}{}$ instruction (the $o$ refers to the seal it uses, namely the offset in the seals made available by linking).

@@

\begin{definition}[Programs and Contexts]
  We define a program as a $(\var{comp},c_{\mathrm{main}})$ component with an empty import list and $c_{\mathrm{main}}$ is one of the exported symbols.

  Given a component $\var{comp}$, a context is a tuple $(\var{comp'},c_{\mathrm{main}})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'},c_{\mathrm{main}})$ is a program.
\end{definition}

\section{Compiler}
\label{sec:compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]
\dominique{27-09-2017: Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions? Is there another function that compiles a single instruction to a list of instructions?}
\lau{28-09-2017: Yes, that was the thought at the moment, but this may very well change soon. This compiler would then use a compiler that compiles single instructions. That said, this part was added a long time ago and for some reason (some details about the compilation probably did not make sense to me at the moment), I did not complete it. I think the compiler will probably change to be from programs to programs, and a programs is not just going to be a list of instructions.}

>>>>

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \ints \\
\com{r} &\in &  \tRegName \\
\com{\rn} &\defbnf &  \com{r} \mid n \\
\com{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \\
      & &  \tgete{\com{r}}{\com{r}}\mid \tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tstore{\com{r}}{\com{r}} \mid\\
      & &  \tload{\com{r}}{\com{r}} \mid \tcca{\com{r}}{\com{\rn}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \\
  & & \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tsetatob{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid \tcseal{\com{r}}{\com{r}} \mid \\ 
      & &   \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\tsplice{\com{r}}{\com{r}}{\com{r}} 
\end{array}
\]
$i$ defines the set $\Instr$

@@

The source machine is also a capability machine with memory capabilities and sealed capabilities. Unlike the target machine, the source machine has a built in stack along with special stack and return tokens used in place of the actual capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    o & \in & \nats \\
    \src{i} & \defbnf &  \com{i} \mid \scall{\offpc,\offsigma}{r}{r}
  \end{array}
\]
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{\offpc,\offsigma}{}{}$ instruction (the $o$ refers to the seal it uses, namely the offset in the seals made available by linking).

@@

\begin{definition}[Programs and Contexts]
  We define a program as a $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ component with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.

  Given a component $\var{comp}$, a context is a tuple $(\var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is a program.
\end{definition}

\begin{definition}[Initial execution configuration]
  \begin{mathpar}
    \inference{
      \reg(\pc) = c_{\mathrm{main},c} & 
      \reg(\rdata) = c_{\mathrm{main},d} \\
      \src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
      \trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
      \overline{c_{\mathrm{import}}} = \emptyset
    }{
      ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
    }
  \end{mathpar}
\end{definition}

\section{Compiler}
\label{sec:compiler}
\[
\comp{\cdot} : \src{i}^* \fun \com{i}^*
\]
\dominique{27-09-2017: Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions? Is there another function that compiles a single instruction to a list of instructions?}
\lau{28-09-2017: Yes, that was the thought at the moment, but this may very well change soon. This compiler would then use a compiler that compiles single instructions. That said, this part was added a long time ago and for some reason (some details about the compilation probably did not make sense to me at the moment), I did not complete it. I think the compiler will probably change to be from programs to programs, and a programs is not just going to be a list of instructions.}
